A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-02 paper/paper.tex
monsterissimo pass

\section{Reinforcement learning: A process model for DMN function}
The reviewed neurobiological properties and likely functions of the DMN
are now argued to be sufficient
for implementing all the major components 
of a full-fledged \textit{reinforcement learning (RL)} algorithm.
Memory, reward signaling,
information gathering, random sampling, and other components
of intelligent RL agents appear to cross-talk in the DMN in a plausible way.

RL is a problem-solving technique in which, through their interactions with an external environment,
an agent learns to reach goals, optimize reward signals, or minimize costs
in an iterative trial-by-error fashion.
At any moment, each taken \textit{action} $a$ triggers a changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward should be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can increase over time.
The environment is generally taken as stochastic,
that is, unpredictable, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
This unpredictability of the environment
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that the more the external world is predictable,
the more mental activity becomes detached from the actual sensory inputs
\citep{antrobus1966studies, pope1978regulation},
the more stimulus-independent thoughts occur, and
the higher the neural activity in the DMN.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to
optimizing control of the organism.
Such a framework of DMN funtion based on reinforcement learning
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of
particularly advanced cognitive processes
can be parsimoneously explained as prediction error minimization
coupled with probabilistic simulations in form of mental scenes,
thus calibrating action portfolios and
maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
If the process has $memory$ such that the next state depends not only on the current state but also on some finite
number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular computational model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP formalizes a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., is likely to be different for another agent in the same state).
This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.

Model-free RL is plausible in the human brain. Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
  is a rather direct mapping of model-free RL learning algorithms
onto the brain,
with the neuromodulator dopamine serving
as a teaching signal to train values or policies by controlling
synaptic plasticity at targets such as the ventral and
dorsolateral striatum.

In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
For instance, if a rat in a maze knows
that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly myopic agent who is solely concerned about its
immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Such an agent is thus better emphasizing consideration of rewards expected in
a short future and paying less attention to
rewards expected at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can then be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

\subsection{The Q-value: A goodness measure for action choices}
The purpose of the DMN can then be
formulated as finding a policy $\pi$ for the agent which maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ thereafter to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
When a choice of actions is available, the $Q^{\pi}(s,a)$
offers a formalization of optimal behavior,
potentially driven by the DMN.

\subsection{Computing optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
This is due to the enormous number of possible action
trajectories the agent might need to
consider simultaneously.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular solution concept is called \text{Q-learning}. It remedies
the issue by only considering the sub-family of
\textit{deterministic policies}, which map each state to the single best action to
take at the state. Such policies define deterministic functions from states to actions.

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), generates a
\textit{greedy policy}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
  \end{split}
  \label{eq:bellman}
\end{equation}
This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function into the immediate reward
and the discounted rewards of the upcoming states.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman}, the sampling is done only over things which
depend on the environment, and so can be learned off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

@@

Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k+1})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
  \label{eq:oracle}
\end{equation}
where
$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
For instance, a general linear model with a kernel $\phi$ would be of the
form
$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation in an
appropriate form of the state-action pairs $(s,a)$,
as proposed by \cite{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed, practically successful alternative
\citep{mnih2015,silver2016mastering} is to learn this
representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
are by far the current state-of-the-art in RL.
The set of $\theta^Q$ parameters that instatiate the nonlinear interactions
between layers of the artifical neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

Learning
of the the entire model parameters can effectively be achieved via \textit{backprop},
wherein prediction errors (i.e., \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

@@

\subsubsection{Link to classical reinforcement learning algorithms} One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
cast in this general variational framework. For instance, TD corresponds to
the above framework using a linear value approximator with feature encoding
$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
$\mathcal S \times \mathcal A$, and so
$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
$$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y),$$
the well-known TD update rule \citep{sutton1998reinforcement}.

@@

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017}.
% These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics. This would be a simple control algorithm by evaluating all legal actions and selecting the action with
% \textbf{highest expected cumulative rewards [please double-check]}.
In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence.} For sampling-based methods it is still almost non-existent beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

\paragraph{Hippocampal replay equals inverse reinforcement-learning?}
%\begin{mdframed}
  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
  For instance, given traces of motor actions from a an adult to grab a cup on a table,
  can an observing child figure out what ``langragian'' functional is being minimized by the former?
  How? Can they reproduce this optimal behavior?
%\end{mdframed}
Such questions are of course pertinent to our decision-making theory for the DMN.
In the general artificial case, the problem has been extensively studied and partially solved by
\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name
``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...

@@

\subsection{Putting everything together}
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
  \caption{The DMN as an RL agent... \textbf{Please add a figure caption}}
  \label{fig:rl_process_chart}
\end{figure}
The DMN is today known to consistently increase in neural
  activity when humans engage in goal-directed behavior that are detached from
  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
  and it was proposed to be situated at the top of the functional network hierarchies
  \citep{carhart2010default, margulies2016situating}.
  Its involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the implicit computation of
  action and state cascades as a function of what happened in the past.
% A policy represents the repertoire of possible actions
%   on the world given a current state. It encodes the probabilites of
%   choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
%   information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and their
  cumulative outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of mind-wandering
  in humans.

The HC may provide perturbed action-transition-state-reward samples
  as a block of "imagined", "hypothesized", "recalled" experience.
  The small variations in these experience blocks allow searching
  a larger space of model parameters and possible experiences.
  Taken to its extreme, stochastic recombination of experience
  building blocks can further optimize the behavior of the RL agent
  by model learning from scenarios in the environment that the agent might
  only very rarely or never encounter.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing (i.e., emulating) possible
  future scenarios and action outcomes.
  Our approach acknowledges the unavoidable stochasticity of
  computation in neuronal processes \citep{faisal2008noise}.

>>>>

\section{Reinforcement learning: A process model for DMN function}
The outlined neurobiological properties and suspected functions of the DMN
are argued to be sufficient
for implementing all components 
of a full-fledged \textit{reinforcement learning (RL)} algorithm.
Integration of past experience, probabilistic mapping of candidate actions,
random sampling of possible scenarios as well as
incorporation of instantaneous, delayed, and expected reward outcomes
are key components of intelligent RL agents
that are plausible to cross-talk in the DMN.

RL is a problem-solving technique in which, through their interactions with an external environment,
an agent learns to reach goals, optimize reward signals, or minimize costs
in an iterative trial-by-error fashion.
At a given moment, each taken \textit{action} $a$ triggers a changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
$r = r(s, a)$ (or regret, since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can grow with time.
The environment is generally taken as stochastic,
that is, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
The unpredictability of the environment
is realistic in a computational model which sets out
to explain DMN function. Indeed, it is known that the more the external world is predictable,
the more mental activity becomes detached from the actual sensory inputs
\citep{antrobus1966studies, pope1978regulation},
the more stimulus-independent thoughts occur, and
the higher the neural activity in the DMN.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to
optimizing control of the organism.
Such a framework of DMN funtion based on reinforcement learning
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of
particularly advanced cognitive processes
can be parsimoneously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
to calibrate action portfolios and
maximize reward outcome at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space \textbf{Replace by "state space"?}.
The process has $memory$ if the next state depends not only on the current state but also on some finite
number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular computational model for
multi-step decision processes in such an environment are
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP formalizes a sequential decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., is likely to be different for another agent in the same state).
This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL},
which we propose as a useful formal framework for DMN function.

Model-free RL is plausible in the human brain. Indeed,
it has long been proposed \citep{dayan2008decision} that there
is a rather direct mapping of model-free RL learning algorithms
onto the brain,
with the neuromodulator dopamine serving
as a teaching signal to train values or policies by controlling
synaptic plasticity at targets such as the ventral and
dorsolateral striatum.
In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
For instance, if a rat in a maze knows
that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

\subsection{The Q-value: A goodness measure for action choices}
We can now formulate the purpose of the DMN
as finding a policy $\pi$ for the human agent that maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
When a choice of actions is available, the $Q^{\pi}(s,a)$
offers a formalization of optimal behavior,
potentially driven by the DMN in human agents.

\subsection{Efficient computation of optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
This is due to the enormous number of possible action
trajectories the agent might need to
consider simultaneously.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular solution is called \text{Q-learning}. It remedies
the issue by only considering the sub-family of
\textit{deterministic policies}, which map each state to the single best action to
take at the state. Such policies define deterministic functions from states to actions.

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
\textit{greedy policy}
\textbf{XXX please add 1-2 sentences why this is interesting in the current
context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
  \end{split}
  \label{eq:bellman}
\end{equation}
This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function \textbf{XXX please more explicitly introduce
the notion of value function}
into the immediate reward
and the discounted rewards of the upcoming states.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman} the sampling is done only over things which
depend on the environment, and so can be learned off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

@@

Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k+1})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
  \label{eq:oracle}
\end{equation}
where
$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
For instance, a general linear model with a kernel $\phi$ would be of the
form
$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation in an
appropriate form of the state-action pairs $(s,a)$,
as proposed by \citep{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed, practically successful alternative approach
\citep{mnih2015,silver2016mastering} is to learn this
representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
likely are state-of-the-art in RL.
The set of $\theta^Q$ parameters that instantiate the nonlinear interactions
between layers of the artifical neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

Learning
of the the entire model parameters can effectively be achieved
via \textit{backpropagation},
where prediction errors (i.e., \textit{regrets}) percolate from lower to higher processing layers to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

@@

\subsubsection{Link to classical reinforcement learning algorithms} One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
cast in this \textbf{please unpack the "this" here}
general variational framework. For instance, TD corresponds to
the above framework using a linear value approximator with feature encoding
$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
$\mathcal S \times \mathcal A$, and so
$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
$$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y),$$
the well-known TD update rule \citep{sutton1998reinforcement}.

@@

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
observations and rewards \citep{aronov2017}.
These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics. This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX Elvis: refs needed in last sentence.} For sampling-based methods it is still almost non-existent beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}

\paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
%\begin{mdframed}
  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
  For instance, given traces of motor actions from an adult to grab a cup on a table,
  can an observing child figure out what ``langragian'' functional is being minimized by the former?
  How? Can they reproduce this optimal behavior?
%\end{mdframed}
Such questions are of course pertinent to our decision-making theory for the DMN.
In the general artificial case, the problem has been extensively studied and partially solved by
\cite{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...

@@

\subsection{Putting everything together}
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
  \caption{The DMN as an RL agent... \textbf{Please add a figure caption}}
  \label{fig:rl_process_chart}
\end{figure}
The DMN is today known to consistently increase in neural
  activity when humans engage in goal-directed behavior that are detached from
  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
  and it was proposed to be situated at the top of the functional network hierarchies
  \citep{carhart2010default, margulies2016situating}.
  Its involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy represents the repertoire of possible actions
    on the world given a current state. It encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
%   information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of mind-wandering
  in humans.

The HC may provide perturbed action-transition-state-reward samples
  as batches of "imagined", "hypothesized", "recalled" experience.
  The small variations in these experience samplings allow searching
  a larger space of model parameters and possible experiences.
  Taken to its extreme, stochastic recombination of experience
  building blocks can further optimize the behavior of the RL agent
  by model learning from scenarios in the environment that the agent might
  only very rarely or never encounter.
  An explanation is thus offered for encountering seemingly familiar situations that
  a human has however never actually encountered (i.e., d\'{e}j\`{a} vu effect).
  While such a situation may not have been experienced in the physical world,
  the DMN may have already stochastically generated, evaluated, and adapted to
  such a randomly synthesized situation in the past.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing (i.e., emulating) possible
  future scenarios and action outcomes.
  Our approach thus acknowledges the unavoidable stochasticity of
  computation in neuronal processes \citep{faisal2008noise}.
